---
title: "Digeorge"
output: html_document
---

The goal of this analysis is to mine EHR data for prevalence of HPO terms in DiGeorge patients with Psychosis in a case control format


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
pacman::p_load(tidyverse, bigrquery, dbplyr)
```



This is how to get the dogeorge patients 
```{r}
cohorts %>% 
  filter(cohort == "digeorge")
```

Setting the connection to the lab
```{r}
bq_auth(path='/vault/secrets/gcp-sa-default.json')  # this path will change to the env variable, but showing for now
project_id <- bq_projects() 
```

```{r}
library(DBI)
con <- dbConnect(
  bigrquery::bigquery(),
  project = "scit414-psd-ae773671", 
  dataset = "arcus"
)
```


```{r}
d <- tbl(con, in_schema("arcus", "diagnosis_icd10"))

d %>% 
  collect()
```


```{r}
DBI::dbGetQuery(con, "Select epic_dx_id 
From arcus.diagnosis_icd10 
Limit 10")
```

##Using ICD codes to identify DiGeorge patients with psychosis 

```{sql, connection = con, output.var = "dg_psyc"}

With dx10 as (
Select epic_dx_id 
From arcus.diagnosis_icd10
Where code in (
"F06.0", 
"F06.1", 
"F06.2",
"F10.150", 
"F10.151", 
"F10.159", 
"F10.250", 
"F10.251", 
"F10.259", 
"F10.950", 
"F10.951", 
"F10.959", 
"F11.150", 
"F11.151", 
"F11.159", 
"F11.250", 
"F11.251", 
"F11.259", 
"F11.950", 
"F11.951", 
"F11.959", 
"F12.150", 
"F12.151", 
"F12.159", 
"F12.250", 
"F12.251", 
"F12.259", 
"F12.950", 
"F12.951", 
"F12.959", 
"F13.150", 
"F13.151", 
"F13.159", 
"F13.250", 
"F13.251", 
"F13.259", 
"F13.950", 
"F13.951", 
"F13.959", 
"F14.150", 
"F14.151", 
"F14.159", 
"F14.250", 
"F14.251", 
"F14.259", 
"F14.950",
"F14.951", 
"F14.959", 
"F15.150", 
"F15.151", 
"F15.159", 
"F15.250", 
"F15.251", 
"F15.259", 
"F15.950", 
"F15.951", 
"F15.959", 
"F16.150", 
"F16.151", 
"F16.159", 
"F16.250", 
"F16.251", 
"F16.259", 
"F16.950", 
"F16.951", 
"F16.959", 
"F17.150",
"F17.151", 
"F17.159", 
"F17.250", 
"F17.251", 
"F17.259", 
"F17.950", 
"F17.951", 
"F17.959", 
"F18.150", 
"F18.151", 
"F18.159", 
"F18.250", 
"F18.251", 
"F18.259", 
"F18.950", 
"F18.951",
"F18.959", 
"F19.150", 
"F19.151", 
"F19.159", 
"F19.250", 
"F19.251", 
"F19.259", 
"F19.950", 
"F19.951", 
"F19.959", 
"F20.0", 
"F20.1", 
"F20.2", 
"F20.3", 
"F20.5", 
"F20.89", 
"F20.9", 
"F20.81", 
"F22", 
"F23", 
"F24", 
"F25.0", 
"F25.1", 
"F25.", 
"F25.8", 
"F25.9", 
"F28", 
"F29", 
"F30.2", 
"F31.2", 
"F31.5", 
"F32.3", 
"F33.3", 
"F53", 
"R44.0", 
"R44.1", 
"R44.2", 
"R44.3", 
"R44.8", 
"R44.9"
)),
dx9 as (

Select epic_dx_id
From arcus.diagnosis_icd9
Where code in (
"293.81",
"293.82", 
"291.3", 
"291.5", 
"292.11", 
"292.12", 
"291.5", 
"291.3", 
"295.00", 
"295.01", 
"295.02", 
"295.03", 
"295.04",
"295.10", 
"295.11", 
"295.12", 
"295.13", 
"295.14", 
"295.15", 
"295.20", 
"295.21", 
"295.22", 
"295.23", 
"295.24", 
"295.25",
"295.30", 
"295.31", 
"295.32", 
"295.33", 
"295.34", 
"295.35", 
"295.50", 
"295.51", 
"295.52", 
"295.53", 
"295.54", 
"295.55", 
"295.60", 
"295.61", 
"295.62", 
"295.63", 
"295.64", 
"295.65", 
"295.80", 
"295.81", 
"295.82", 
"295.83", 
"295.84", 
"295.85", 
"295.90", 
"295.91", 
"295.92", 
"295.93", 
"295.94", 
"295.95", 
"295.40", 
"295.41", 
"295.42", 
"295.43", 
"295.44", 
"295.45", 
"297.0", 
"297.1", 
"297.2", 
"298.3", 
"298.4", 
"297.3", 
"295.70", 
"295.71", 
"295.72", 
"295.73", 
"295.74", 
"295.75", 
"298.8", 
"298.9", 
"296.04", 
"296.44", 
"296.54", 
"296.24", 
"298.0", 
"296.34", 
"780.1"
)),

dx as (
Select epic_dx_id From dx9 
union ALL
select epic_dx_id from dx10 
),


en as (
Select d.epic_pat_id, d.epic_dx_id 
From arcus.encounter_diagnosis as d 
Join dx
on d.epic_dx_id = dx.epic_dx_id 
),

co as (
Select epic_pat_id
From arcus.cohorts
Where cohort = "digeorge"
)

Select co.epic_pat_id 
From co 
Join en
on en.epic_pat_id = co.epic_pat_id

```


```{sql, connection = con, output.var = "dg_appt"}
With c as (
   Select Distinct  epic_pat_id
   From arcus.cohorts
   Where cohort = "digeorge"
)


Select Distinct p.epic_pat_id, appt_age
From arcus.encounter as p 
Join c on c.epic_pat_id = p.epic_pat_id
Where appt_age is not null
Order by epic_pat_id

```


ICD 10

```{sql, connection = con, output.var = "psychosis"}


With dx10 as (
Select epic_dx_id 
From arcus.diagnosis_icd10
Where code in (
"F06.0", 
"F06.1", 
"F06.2",
"F10.150", 
"F10.151", 
"F10.159", 
"F10.250", 
"F10.251", 
"F10.259", 
"F10.950", 
"F10.951", 
"F10.959", 
"F11.150", 
"F11.151", 
"F11.159", 
"F11.250", 
"F11.251", 
"F11.259", 
"F11.950", 
"F11.951", 
"F11.959", 
"F12.150", 
"F12.151", 
"F12.159", 
"F12.250", 
"F12.251", 
"F12.259", 
"F12.950", 
"F12.951", 
"F12.959", 
"F13.150", 
"F13.151", 
"F13.159", 
"F13.250", 
"F13.251", 
"F13.259", 
"F13.950", 
"F13.951", 
"F13.959", 
"F14.150", 
"F14.151", 
"F14.159", 
"F14.250", 
"F14.251", 
"F14.259", 
"F14.950",
"F14.951", 
"F14.959", 
"F15.150", 
"F15.151", 
"F15.159", 
"F15.250", 
"F15.251", 
"F15.259", 
"F15.950", 
"F15.951", 
"F15.959", 
"F16.150", 
"F16.151", 
"F16.159", 
"F16.250", 
"F16.251", 
"F16.259", 
"F16.950", 
"F16.951", 
"F16.959", 
"F17.150",
"F17.151", 
"F17.159", 
"F17.250", 
"F17.251", 
"F17.259", 
"F17.950", 
"F17.951", 
"F17.959", 
"F18.150", 
"F18.151", 
"F18.159", 
"F18.250", 
"F18.251", 
"F18.259", 
"F18.950", 
"F18.951",
"F18.959", 
"F19.150", 
"F19.151", 
"F19.159", 
"F19.250", 
"F19.251", 
"F19.259", 
"F19.950", 
"F19.951", 
"F19.959", 
"F20.0", 
"F20.1", 
"F20.2", 
"F20.3", 
"F20.5", 
"F20.89", 
"F20.9", 
"F20.81", 
"F22", 
"F23", 
"F24", 
"F25.0", 
"F25.1", 
"F25.", 
"F25.8", 
"F25.9", 
"F28", 
"F29", 
"F30.2", 
"F31.2", 
"F31.5", 
"F32.3", 
"F33.3", 
"F53", 
"R44.0", 
"R44.1", 
"R44.2", 
"R44.3", 
"R44.8", 
"R44.9"
)),
dx9 as (

Select epic_dx_id
From arcus.diagnosis_icd9
Where code in (
"293.81",
"293.82", 
"291.3", 
"291.5", 
"292.11", 
"292.12", 
"291.5", 
"291.3", 
"295.00", 
"295.01", 
"295.02", 
"295.03", 
"295.04",
"295.10", 
"295.11", 
"295.12", 
"295.13", 
"295.14", 
"295.15", 
"295.20", 
"295.21", 
"295.22", 
"295.23", 
"295.24", 
"295.25",
"295.30", 
"295.31", 
"295.32", 
"295.33", 
"295.34", 
"295.35", 
"295.50", 
"295.51", 
"295.52", 
"295.53", 
"295.54", 
"295.55", 
"295.60", 
"295.61", 
"295.62", 
"295.63", 
"295.64", 
"295.65", 
"295.80", 
"295.81", 
"295.82", 
"295.83", 
"295.84", 
"295.85", 
"295.90", 
"295.91", 
"295.92", 
"295.93", 
"295.94", 
"295.95", 
"295.40", 
"295.41", 
"295.42", 
"295.43", 
"295.44", 
"295.45", 
"297.0", 
"297.1", 
"297.2", 
"298.3", 
"298.4", 
"297.3", 
"295.70", 
"295.71", 
"295.72", 
"295.73", 
"295.74", 
"295.75", 
"298.8", 
"298.9", 
"296.04", 
"296.44", 
"296.54", 
"296.24", 
"298.0", 
"296.34", 
"780.1"
)),

dx as (
Select epic_dx_id From dx9 
union ALL
select epic_dx_id from dx10 
), 


d as(
Select Distinct d.epic_pat_id, d.epic_enc_id 
From arcus.encounter_diagnosis as d 
Join dx
on d.epic_dx_id = dx.epic_dx_id 
)

Select d.epic_pat_id, appt_age 
From arcus.encounter as e
Join d
on d.epic_enc_id = e.epic_enc_id
Where appt_age is not null
```






```{r}
cohorts %>% 
  filter(cohort == "digeorge") %>% 
  inner_join(dg_psyc, by = "epic_pat_id") %>% 
  distinct()
```


```{r}
distinct(dg_psyc)

cohorts %>% 
  filter(cohort == "digeorge") %>% 
  distinct

head(df.ages)

```

Appt ages 22q patients

```{r}
dg_appt_max_year <- dg_appt %>% 
  group_by(epic_pat_id) %>% 
  summarise(appt_age = max(appt_age), .groups = "drop") %>% 
  mutate(age_year = appt_age/365) %>% 
  select(-appt_age)
```



```{r}
psychosis_min_year <- psychosis %>% 
  group_by(epic_pat_id) %>% 
  summarise(appt_age = min(appt_age), .groups = "drop") %>% 
  mutate(age_year = appt_age/365) %>% 
  select(-appt_age)
```



```{r}
all_age_dg_p <- dg_appt_max_year %>% 
  full_join(psychosis_min_year, by = "epic_pat_id") %>% 
  mutate(age_year = coalesce(age_year.x, age_year.y)) %>% 
  mutate(disease  = case_when(is.na(age_year.x) ~ "Psychosis", 
                   is.na(age_year.y) ~ "Digeorge", 
                   TRUE              ~ "Both"
    
  ))


```



```{r}
all_age_dg_p %>% 
  group_by(epic_pat_id) %>% 
  ggplot(aes(x = age_year, col = disease))+
  geom_density()

```



```{r}
 dg_icd_hpo <-cohorts %>% 
  filter(cohort == "digeorge") %>% 
  inner_join(dg_psyc, by = "epic_pat_id") %>% 
  distinct() %>% 
  select(epic_pat_id) %>% 
  inner_join(note_info %>% 
               select(epic_pat_id, epic_note_id) %>% 
               distinct(), by = "epic_pat_id") %>% 
  inner_join(cohort_annotations %>% 
             select(ontology_name, ontology_code, ontology_preferred_text, epic_note_id), 
             by = "epic_note_id")

```


```{r}
dg_icd_hpo %>% 
  group_by(ontology_preferred_text) %>% 
  summarise(unique_pat = n_distinct(epic_pat_id), .groups = "drop") %>% 
  arrange(desc(unique_pat))
```


#Data set - first time that the 22q patients identified by HPO came in with psychosis 
```{r}

dgp_hpo <- cohorts %>%
  filter(cohort == "digeorge") %>%
  select(cohort, epic_pat_id) %>% 
  distinct() %>% 
  inner_join(note_info %>%
               select(epic_pat_id, epic_note_id) %>%
               distinct(),
             by = "epic_pat_id") %>%
  distinct() %>% 
  inner_join(
    cohort_annotations %>%
      select(ontology_preferred_text, epic_note_id) %>%
      filter(ontology_preferred_text == "Psychotic Disorders") %>% 
      distinct(),
    by ="epic_note_id"
 ) %>% 
  inner_join(note_info %>% select(epic_note_id, epic_enc_id, note_entry_datetime) %>% 
               distinct(), 
             by = "epic_note_id") %>% 
  group_by(epic_pat_id) %>% 
  arrange(note_entry_datetime) %>%
  slice(1) %>% 
  ungroup()

head(dgp_hpo)

dgp_hpo %>% 
  summarise(n_distinct(epic_pat_id))
  
```

#Looking at the ages of the 63 22q patients' first encounter with "Psychotic Disorders" as identified by HPO
```{r}
head(dgp_hpo_age)

dgp_hpo_age <- dgp_hpo %>% 
  inner_join(encounter %>% 
               select(epic_pat_id, epic_enc_id, appt_age, contact_age, effective_age) %>% 
               distinct(), 
             by = c("epic_pat_id", "epic_enc_id")) %>% 
  #na.omit() %>%
  mutate(age_year = coalesce(appt_age, contact_age, effective_age)/365)


head(dgp_hpo_age)

```


# Density plot of ages at first encounter of the digeorge psychotic patients. 
```{r}
dgp_hpo_age %>% 
  ggplot(aes(x = age_year))+
  geom_density()
  
```


# Getting a count of how many patients I have for each age year.

```{r}
dgp_hpo_age_count <- dgp_hpo_age %>% 
  mutate(age_year_r = floor(age_year)) %>% 
  group_by(age_year_r) %>% 
  summarise(count = n(), .groups= "drop") 



dgp_hpo_age_count 
  
```
```{r}
dgp_hpo_age %>% 
  filter(floor(age_year) == -1) %>% 
  slice(1) %>% 
  select(epic_pat_id) %>% 
  inner_join(encounter, by = "epic_pat_id")
```










#getting a data set of 22q w/o psychosis - determination by HPO and age matching 

```{r}

dg_np_hpo <- cohorts %>%                               ## did an antijoin with the dgp_hpo data set of 63 patients to get np 
  filter(cohort == "digeorge") %>%
  select(cohort, epic_pat_id) %>% 
  distinct() %>% 
  anti_join(dgp_hpo, by = "epic_pat_id") 



##ages

dg_np_age_year_hpo <- dg_np_hpo%>% 
   inner_join(encounter %>% 
               select(epic_pat_id, epic_enc_id, appt_age, effective_age, contact_age) %>% 
               distinct(), 
             by = c("epic_pat_id")) %>% 
  #na.omit() %>%
  mutate(age_year = coalesce(appt_age, effective_age, contact_age)/365) %>% 
  mutate(age_year_r = floor(age_year))





```

Counts

```{r}

dg_np_age_year_hpo_20 <- dg_np_age_year_hpo %>%   #Taking just people 20 or older
  filter(age_year_r > 20) %>% 
  group_by(epic_pat_id) %>% 
  sample_n(1) %>% 
  ungroup()

dg_np_age_year_hpo_18_20 <- dg_np_age_year_hpo %>%     # Taking people between 18-20
  filter(age_year_r > 18, age_year_r <= 20) %>% 
  group_by(epic_pat_id) %>% 
  sample_n(1) %>% 
  ungroup() 

dgnp_hpo_age_year_1 <-dg_np_age_year_hpo %>% 
  anti_join(dg_np_age_year_hpo_18, by = "epic_pat_id") %>%  #anti-Joining the two 
  group_by(epic_pat_id) %>% 
  sample_n(1) %>% 
  ungroup() %>% 
  bind_rows(dg_np_age_year_hpo_18, dg_np_age_year_hpo_18_20) 

dgnp_hpo_age_count <- dgnp_hpo_age_year_1 %>%      #
  mutate(age_year_r = floor(age_year)) %>% 
  group_by(age_year_r) %>% 
  summarise(count_np = n_distinct(epic_pat_id), .groups= "drop") 

dgp_hpo_age_count <- dgp_hpo_age %>% 
  mutate(age_year_r = floor(age_year)) %>% 
  group_by(age_year_r) %>% 
  summarise(count_p = n(), .groups= "drop") 


dgp_hpo_age_count %>% 
  full_join(dgnp_hpo_age_count, by = "age_year_r") %>%   #Testing to see if we have at least one more np patient for every p 
  arrange(age_year_r) %>% 
  filter(!is.na(count_p))





  



```






# CASE CONTROL MATCHING
```{r}
control_cohort <- dgnp_hpo_age_year_1 %>% 
  select(epic_pat_id, age_year_r) %>% 
  distinct() %>% 
  inner_join(dgp_hpo_age_count, by = "age_year_r") %>% 
  group_by(age_year_r) %>%
  sample_n(count_p) %>% 
  ungroup() %>% 
  arrange(epic_pat_id) 

control_cohort %>% 
  head(10)




###These are the non-psychotic patient id's 




```


##Adding notes to the data set

```{r}
control_cohort_enc <- dg_np_age_year_hpo %>% 
  mutate(age_year_r = floor(age_year)) %>% 
  inner_join(control_cohort, by = c("epic_pat_id", "age_year_r")) %>% 
  group_by(epic_pat_id) %>% 
 # slice(1:10) %>% 
  ungroup() 
head(control_cohort_enc)

```


```{r}
c_n <- control_cohort_enc %>% 
  inner_join(note_info, by = c("epic_pat_id", "epic_enc_id")) %>% 
  group_by(epic_pat_id) %>% 
  slice(1:4) %>% 
  ungroup() %>% 
  select(cohort, epic_pat_id, epic_enc_id, epic_note_id, age_year_r, note_entry_datetime)
```

```{r}
c_n %>% 
  summarise(n_distinct(epic_note_id))
```
```{r}
np_notes_age_match <-c_n %>% 
  inner_join(dg_np_hpo_codes, by = c("epic_pat_id","epic_note_id")) %>% 
  mutate(case = 1)

head(np_notes_age_match)
```

```{r}
p_notes_age_match <- dgp_hpo_age %>% 
  select(cohort, epic_pat_id, epic_enc_id, epic_note_id, age_year, note_entry_datetime) %>% 
  inner_join(dgp_hpo_codes, by = c("epic_pat_id", "epic_note_id")) %>% 
  mutate(age_year_r = round(age_year)) %>% 
  select(-age_year) %>% 
  mutate(case=0)

p_notes_age_match

```


```{r}
p_np_age_match_merged<- np_notes_age_match %>% 
  bind_rows(p_notes_age_match) 
```

```{r}
p_np_age_match_merged %>% 
  group_by(case) %>% 
  summarise(n_distinct(epic_pat_id), .groups ="drop") 

```




```{r}
control_cohort %>% 
  group_by(epic_pat_id) %>% 
  summarise(n()) 
```








```{r}
dgp_hpo_codes <- dgp_hpo %>% 
  select(epic_pat_id) %>% 
  distinct() %>% 
  inner_join(note_info, by = "epic_pat_id") %>% 
  inner_join(cohort_annotations, by = "epic_note_id") %>% 
  select(epic_pat_id, epic_note_id, ontology_code, ontology_preferred_text) %>%
  filter(ontology_preferred_text != "Psychotic Disorders")

head(dgp_hpo_codes)
  
  
```

```{r}
dg_np_hpo_codes <- dg_np_hpo %>% 
  select(epic_pat_id) %>% 
  distinct() %>% 
  inner_join(note_info, by = "epic_pat_id") %>% 
  inner_join(cohort_annotations, by = "epic_note_id") %>% 
  select(epic_pat_id, epic_note_id, ontology_code, ontology_preferred_text)
 # filter(ontology_preferred_text != "Psychotic Disorders")

head(dg_np_hpo_codes)
  
```









Doing a similarity analysis 

```{r}
library(ontologyIndex)
library(ontologySimilarity)
data(hpo)
```

```{r}
ontologies <-tibble(ontology_code= hpo$id)

head(ontologies)
```


```{r}
term_sets_H<- dgp_hpo_codes %>% 
 # filter(epic_pat_id != "Z101515") %>% 
  inner_join(ontologies, by = "ontology_code") %>% 
  #slice(1:2000) %>% 
  select(epic_pat_id, ontology_code) %>% 
  group_by(epic_pat_id) %>% 
  slice(1:10) %>% 
  summarise(ID = list(ontology_code), .groups = "drop") %>% 
  select(ID) %>% 
  pull(ID)

```



```{r}
sim_mat <- get_sim_grid(ontology=hpo, term_sets=term_sets_H)
```


```{r}
get_sim(sim_mat, group = 1:63)
```

```{r}
get_sim_p(sim_mat, group=c(1:63))
```


```{r}
dist_mat <- max(sim_mat) - sim_mat
plot(hclust(as.dist(dist_mat)))
```





```{r}
head(dgp_hpo_codes)
head(dg_np_hpo_codes)

```



```{r}
dg_np_p_hpo <- dgp_hpo_codes %>% 
  bind_rows(dg_np_hpo_codes)

head(dg_np_p_hpo)
```


```{r}
ontologies <-tibble(ontology_code= hpo$id)

head(ontologies)
```




```{r}
term_sets_H_two<- dg_np_p_hpo %>% 
 # filter(epic_pat_id != "Z101515") %>% 
  inner_join(ontologies, by = "ontology_code") %>% 
  #slice(1:2000) %>% 
  select(epic_pat_id, ontology_code) %>% 
  group_by(epic_pat_id) %>% 
  slice(1:10) %>% 
  summarise(ID = list(ontology_code), .groups = "drop") %>% 
  select(ID) %>% 
  pull(ID)
```

```{r}
sim_mat_2 <- get_sim_grid(ontology=hpo, term_sets=term_sets_H_two)
```



```{r}
get_sim(sim_mat_2, group = 1:1008)
```




```{r}
get_sim_p(sim_mat_2, group=c(1:1008))
```
```{r}
dist_mat <- max(sim_mat_2) - sim_mat_2
plot(hclust(as.dist(dist_mat)))
```

```{r}
h_1 <- hclust(as.dist(dist_mat))


c1 <- cutree(h_1, k =2)

plot(c1)
```



















