---
title: "gamm_interaction"
author: "Harshini Raman"
date: '2022-11-19'
output: html_document
---


# Loading libraries, setting plot parameters 

# case = psychosis
# non-psychotic = psych illnesses but not psychotic disorders
# healthy = no psych illness

## Data sets needed: cohorts, note_info, patient, cohort_annotations  located in arcus > lab > shared > data

  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#install.packages("pacman")  
#install.packages("rlang")
#install.packages("dplyr")
pacman::p_load(tidyverse, dplyr, ggplot2, broom, mgcv)


font_size <- 16
theme_set(theme_classic(base_family = "sans",base_size = font_size))
line_size <- 1.5
point_size <- 2

```


Removing enceph patients from cohorts data set 
```{r}
cohort_type <- cohorts %>%         
  select(epic_pat_id, cohort) 

cohort_type 
```


## Joining information about notes, date of birth, and annotations into the data_for_gam 

```{r}
data_for_gam_long <- cohort_type %>%
  inner_join(note_info %>% select(epic_pat_id, epic_note_id, note_entry_datetime), by = "epic_pat_id")  %>%
  inner_join(patient  %>% select(epic_pat_id, dob), by = "epic_pat_id") %>%
  inner_join(cohort_annotations %>% select(epic_note_id, ontology_preferred_text) %>% distinct() , by = "epic_note_id") %>%
  distinct()

#saveRDS(data_for_gam_long, "arcus/shared/data/data_for_gam_long.rds")  # can save this data set if you want 


data_for_gam_long %>%
  filter(str_detect(ontology_preferred_text, "deficit"))  ## useful if you want to see if an hpo term is in the data set 
  

```
## Pick the HPO term you want to predict 
## one column with the term or "none" if note doesn't have term 

```{r}
data_for_gam_long_dt<- data_for_gam_long %>%
  mutate(ontology_preferred_text = if_else(ontology_preferred_text %in% c("Attention deficit hyperactivity disorder"), ontology_preferred_text, "none")) %>%
  distinct() %>%
  mutate(value = 1) %>%
  data.table::as.data.table()




```
## Making a wide version of the data set
## calculating age in days
## cleaning up the names of the terms 

```{r}
data_for_gam_wide <- data.table::dcast(data = data_for_gam_long_dt, formula = epic_pat_id + cohort + epic_note_id + note_entry_datetime +dob ~ ontology_preferred_text, fill = 0, value.var = "value") %>%
  mutate(age_days = as.integer(lubridate::date(note_entry_datetime) - lubridate::date(dob)))  %>%
  select(-note_entry_datetime, -dob) %>%
  janitor::clean_names() %>%
 filter(!is.na(age_days)) 
```

## Creating a model formula and model object 

```{r}
model_formula <- as.formula("attention_deficit_hyperactivity_disorder ~ s(age_days, k = 3)")
modobj <- gamm(model_formula,
              random = list(epic_pat_id=~1), #We specify a random intercept for subject id (bblid). gamm uses lists for this.
              data = data_for_gam_wide %>% filter(cohort == "case"),  # specify cohort 
              family = binomial, 
              niterPQL=100)        # can change niterPQL (iterations) to how many ever you want 

```

## Plotting our model

```{r}

b_modobj<- b_modobj %>% 
  mutate(age_years = age_days/365) %>%
  filter(!is.na(age_years)) %>%
  mutate(psychotic_disorders = factor(attention_deficit_hyperactivity_disorder, levels = c(1,0), labels = c("ADHD", "Non-ADHD")))


b_modobj %>% 
ggplot(aes(x = age_years, y = exp(.fitted), col = as.factor(attention_deficit_hyperactivity_disorder)))+
geom_jitter(width = 0.5, alpha = 0.5)+
ggtitle("attention_deficit_hyperactivity_disorder") 
```


