---
title: "HPO Sample"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
install.packages("ontologyIndex")
install.packages("ontologySimilarity")
data(hpo)
```

```{r}
library(ontologyIndex)
library(ontologySimilarity)
data(hpo)
set.seed(1)
```

Next, we’ll set the information content for the terms. This is typically based on some kind of ‘population frequency’, for example: the frequency with which the term is used, explicitly or implicity, to annotate objects in a database. Such frequency information is not always available, but it could still be useful to define the information content with respect to the frequency with which the term is an ancestor of other terms in the ontology (as this still captures the structure of the ontology).

```{r}
information_content <- descendants_IC(hpo)
information_content
```

Now we’ll generate some random sets of terms. We’ll sample 5 random term sets (which could for example represent the phenotypes of patients) of 8 terms. Note that here, we call the minimal_set function from the ontologyIndex package on each sample set to remove redundant terms. Typically, ontological annotations would be stored as such minimal sets, however if you are unsure, it is best to call minimal_set on each term set to guarantee the similarity expressions are faithfully evaluated (the package chooses not to map to minimal sets by default for speed).
```{r}
term_sets <- replicate(simplify=FALSE, n=5, expr=minimal_set(hpo, sample(hpo$id, size=8)))
term_sets

typeof(term_sets)
```

Then one can calculate a similarity matrix, containing pairwise term-set similarities:
```{r}
sim_mat <- get_sim_grid(ontology=hpo, term_sets=term_sets)
sim_mat
```
Group similarity of phenotypes 1-3, based on sim_mat:
```{r}
get_sim(sim_mat, group=1:3)
```
p-value for significance of similarity of phenotypes 1-3:
```{r}
get_sim_p(sim_mat, group=1:3)
```

